---
title: "p8105_hw5_sr4081"
author: "Samiha Reza"
date: "2024-11-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
library(ggplot2)
```



## Problem 2

``` {r function()}
n <- 30
sigma <- 5
mu_values <- 0:6
n_datasets <- 5000
alpha <-0.05

set.seed(10)

power_results <- vector ("list", length(mu_values))

for (i in seq_along(mu_values)) {
  mu <- mu_values[i]

results <- replicate(n_datasets, {
    x = rnorm(n, mean = mu, sd = sigma)
    t_test <- t.test(x, mu=0)

    tidy_t_test <- broom::tidy(t_test)
    tidy_t_test$mu <- mu
    tidy_t_test[c("estimate", "p.value", "mu")]
}, simplify = FALSE)

results_df <- do.call(rbind, results)
colnames(results_df) <- c("mu_hat", "p_value", "mu")


power <- mean(results_df$p_value < alpha)


avg_mu_hat <- mean(results_df$mu_hat)
avg_mu_hat_rejected <- mean(results_df$mu_hat[results_df$p_value < alpha])


power_results[[i]] <- data.frame(mu=mu, 
                                 power=power,
                                 avg_mu_hat=avg_mu_hat,
                                 avg_mu_hat_rejected=avg_mu_hat_rejected)

}

final_results_df <- do.call(rbind, power_results)


head(final_results_df)
```
``` {r plot1}

ggplot(final_results_df, aes(x = mu, y = power)) + 
  geom_line(color = "blue") +
  geom_point(color = "red") +
  labs(
    title = "Power vs Effect Size", 
    x = "True mean",
    y = "Power"
  ) + 
  theme_minimal()

```
As the true value of mean increases, the power increases as well. This indicates how effect size and power are directly proportional. 

``` {r plot2}

ggplot(final_results_df, aes (x = mu, y = avg_mu_hat)) + 
  geom_line(aes(y = avg_mu_hat), color = "blue") +
  geom_point(aes(y=avg_mu_hat), color = "blue") +
  
  geom_line(aes(y=avg_mu_hat_rejected), color = "red") +
  geom_point(aes(y=avg_mu_hat_rejected), color = "red") +
  labs(
    title = "Average estimate of mean vs True mean", 
    x = "True mean",
    y = "Average Estimate of Mean"
  ) + 
  theme_minimal()

```
The blue line refers to the average estimate of mean for all generated p-values; the estimates are close to the true mean. The red line refers to the average estimate of mean for p-values that rejected the null hypothesis. The sample average for the rejected null p-values is not approximately equal when the true mean is smaller because of the affect on effect size and power. 

## Problem 3

``` {r data }
url <- "https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"
download.file(url, destfile = "homicide-data.csv")

homicide_data <- read.csv(url) %>%
    janitor::clean_names() %>%
    mutate(city_state = paste(city, state, sep = ", ")
)

homicide_summary <- homicide_data %>%
  mutate(unsolved = ifelse(disposition %in% c("Closed without arrest", "Open/No arrest"), 1, 0)
         ) %>%
  group_by(city_state) %>%
  summarise(
    total_homicide = n(),
    unsolved_homicides = sum(unsolved),
    .groups = "drop"
  )
```
The dataset has `r nrow(homicide_data)` observations and `r ncol(homicide_data)` variables. The variables include `r colnames(homicide_data)`.

